Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-23 src/2.GenericProgramming.tex
With edits via Jan

\begin{figure}[h]
\begin{Verbatim}[frame=leftline]
Constrained Semantics
Parallel Paradigms 
Division of Responsibility
\end{Verbatim}
\caption{Semantic capture - a triplet of constrained semantics, parallel paradigms, and division of responsibilities defined by a parallel programming model to allow developers to run performance-portable code on a set of architectures}
\label{figSemCapture}
\end{figure}

@@

Lastly, a model must divide responsibility between. In practice, this is largely a choicebetween declarative and imperative semantics. Declarative semantics require a developer to express intent, implementation is the responsibility of the programming model and underlying toolchain. Imperative semantics require the developer to specify the mechanisms by which their intent will be implemented. In an imperative model, developer success is tied to their knowledge of implementation mechanisms. In a declarative model that success is tied to the developer's ability to accurately declare their requirements, and the toolchain's ability to turn that into an implementation. Figures~\ref{figOMPLike} and~\ref{figKokkosLike} show examples of two sample applications using the aforementioned types of semantics. While both programming models implement the same programming paradigm (parallel patterns), Figure~\ref{figOMPLike} required the developer to express how to parallelize the loop, while in~\ref{figKokkosLike} the developer just declared their parallel semantics, and relied on the programming model to pick an implementation.

>>>>

\begin{figure}[h]
\begin{Verbatim}[frame=leftline]
Constrained Semantics: memory allocations, memory movement, parallel kernels
Parallel Paradigms: Patterns, Tasks
Declarative vs Imperative Semantics: Declarative
\end{Verbatim}
\caption{Semantic capture - a triplet of constrained semantics, parallel paradigms, and division of responsibilities defined by a parallel programming model to allow developers to run performance-portable code on a set of architectures}
\label{figSemCapture}
\end{figure}

@@

Lastly, a model must divide responsibility between the developer and the model. In practice, this is largely a choice between declarative and imperative semantics. Declarative semantics require a developer to express intent, implementation is the responsibility of the programming model and underlying toolchain. Imperative semantics require the developer to specify the mechanisms by which their intent will be implemented. In an imperative model, developer success is tied to their knowledge of implementation mechanisms. In a declarative model that success is tied to the developer's ability to accurately declare their requirements, and the toolchain's ability to turn that into an implementation. Figures~\ref{figOMPLike} and~\ref{figKokkosLike} show examples of two sample applications using the aforementioned types of semantics. While both programming models implement the same programming paradigm (parallel patterns), Figure~\ref{figOMPLike} required the developer to express how to parallelize the loop, while in~\ref{figKokkosLike} the developer just declared their parallel semantics, and relied on the programming model to pick an implementation.
